Paper 3: Housing and the capital share of income

### 2022-11-30 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
Updating papers and data appendices

```{r}
library(tidyverse)
```

## Introduction

## Raw data files

Make a list of filenames (both to reference the correct file in the url and as the filename to save locally).
```{r}

files<- c("32_countries",
          "Canada",
          "US_alt")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
)

@@

# Add directory to specify filepaths in which to save a local copy of each file
filepaths<- paste0("PhD_pages/data/original/KLEMS/", filenames)

@@

@@

Save files locally.
```{r}

for (file in files){
  print(file)
download.file(download_urls[file], filepaths[file])
}

@@

## Preparing dataframes

Importing the main data file - with data for 32 countries.
```{r}

df <- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df<- df %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df <- df %>% 
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

@@

## Calculatiing derived variables

>>>>

## Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the KLEMS industry data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Load libraries and functions

```{r}
library(tidyverse)
library(readxl) # For handling xls source data files
library(knitr)

source("PhD_pages/scripts/general_functions.R")
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Specify the location of four original data files

We obtain data from four KLEMS datasets:

-   EU KLEMS version ...
-   Canada...
-   An additional series for the US (different industry classification, longer series..)
-   Japan

Here we specify corresponding lists of URLs where the original files are located, short names for the files, and local filepaths where a copy of the original data files will be saved.

```{r}

files<- c("32_countries",
          "CAN",
          "USA")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx"
)

# JAPAN OLD: "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"

@@

# The directory where the copies will be saved
fp<- "PhD_pages/data/original/KLEMS/"

# Append this directory to the file names to specify file paths
filepaths<- paste0(fp, filenames)

@@

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save a local copy of the original data files

If a new copy of the original data is required, the `grab_original` variable can be set to `TRUE`. Otherwise this step is skipped.

```{r}
grab_original<- FALSE

if (grab_original){
  for (file in files){
    print(paste0("Downloading ", file))
    download.file(download_urls[file], filepaths[file])
  }
}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load locally saved copies of original data as data frames

```{r}
df_list<- list()
```

Importing the main data file - with data for 32 countries.

```{r}

df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (Ideally this would be handled by specifying the right format in the read.csv command above. But I was unable to find the right format specification)
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

# Recode the Entity name for US: 'USA-NAICS' -> 'USA'
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = if_else(country == "USA-NAICS",
                           "USA",
                           country))

@@

Data for Canada and the alternative series for the US.

```{r}
countries<- c("CAN", "USA")

for (country in countries){
df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
                                                         sheet = 'DATA'))

# convert to long format and rename variables
df_list[[country]]<- df_list[[country]] %>%
  pivot_longer(cols = starts_with("_"), names_to = "year",
    names_prefix = "_", values_to ="value",
    names_transform = list(year=as.integer)) %>%
  rename(transaction = Variable) %>%
  mutate(country = country)

}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Tidying Canada data

```{r}
# Drop some NA values (- data for 2009 and 2010 is given as missing)
df_list[["CAN"]]<- df_list[["CAN"]] %>%
  filter(!is.na(value))

# For Canada we need to calculate some industry aggregates to be comparable with other countries
agg_industries<- list()

agg_industries[['D']]<- c(
  "15t16",
  "17t19",
  "20",
  "21t22",
  "23",
  "24",
  "25",
  "26",
  "27t28",
  "29",
  "30t33",
  "34t35",
  "36t37")

agg_industries[['G']]<- c(
  "50",
  "52")

agg_industries[['I']]<- c(
  "60t63",
  "64")

agg_industries[['K']]<- c(
  "70",
  "71t74")

for (agg_ind in names(agg_industries)){
  print(paste0("Adding aggregate data for industry: ", agg_ind))

  df_totals<- df_list[["CAN"]] %>%
  filter(code %in% agg_industries[[agg_ind]]) %>%
    group_by(transaction, year) %>%
    summarise(value = sum(value)) %>%
    ungroup() %>%
    mutate(code = agg_ind) %>%
    mutate(country = "CAN")

  df_list[["CAN"]]<- bind_rows(df_list[["CAN"]], df_totals)

}

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Tidying Japan data - OLD, TO DELETE

```{r}
# Drop NA values
# df_list[["JPN"]]<- df_list[["JPN"]] %>% filter(is.na(value))
# 
# # Rename the capital compensation variable, to agree with other files
# df_list[["JPN"]]<- df_list[["JPN"]] %>%
#   mutate(transaction = replace(transaction,
#                                transaction == "CAP_GFCF",
#                                "CAP"))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Join the different dataframes together into a single dataframe

```{r}
df<- bind_rows(
  df_list, .id = "source"
)

# Make a copy of the original merged dataframe, for use later on.
df_orig<- df
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop duplicate series for the US

Data for the US is included in the main data file using the NAICS industry classification.

A separate source provides a much longer series using SIC classification.

Here we show the data for aggregate Value Added from the two datasets.
```{r}

df %>%
  filter(country %in% c("USA"),
         code == "TOT",
 transaction == "VA") %>%
  ggplot(aes(x = year, y = value, shape = source, colour = source, alpha=0.5)) +
  geom_point()

@@

We drop the shorter NAICS data in favour of the SIC data.
```{r}
df<- df %>%
  filter(!(country == "USA" & source == "32_countries"))

@@

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Standardize country names

Here is the mapping of country names we use to standardize across data sources.
```{r}

KLEMS_country_mapping<- read.csv(paste0(fp, "KLEMS_country_name_mapping", ".csv"))

kable(KLEMS_country_mapping)

Here we map the joined KLEMS data sets to these standardized names.
```{r}
df<- map_names(
  df = df, 
  df_mapping = KLEMS_country_mapping, 
  old_country_varname_df = "country", 
  old_country_varname_df_mapping = "Country_KLEMS",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Select relevant transactions and convert to wide format

We keep only the transactions we need: Value Added, Capital income, Labour income and Compensation of employees. And we spread the data to a wider format.

```{r}
df<- df %>%
  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))

df<- df %>%
  pivot_wider(id_cols = c(country, year, code),
              names_from = "transaction",
              values_from = "value")

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Replace some NA values with 0

There are many (`r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()`) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.

Most of these (`r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()`) are where both capital income and value added equal zero. Such observations appear only for the following industries:

```{r}
kable(
  df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count(code) 
)
```

There are also a small number of such observations where VA>0, listed here:

```{r}
kable(
  df %>% filter(VA == CAP & is.na(LAB) & VA>0)
)

For both cases (i.e. where LAB is NA and either VA = CAP = 0 or VA = CAP > 0) we set LAB = 0.

```{r}
df<- df %>%
  mutate(LAB = replace(LAB,
                       VA==CAP & is.na(LAB),
                       0))
```

There is a similar set of country-year-industry observations (n = `r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()`) where VA = LAB but CAP is NA. These we list here.

```{r}
kable(
  df %>% filter(VA == LAB & is.na(CAP) & VA>0)
)

For these cases (i.e. where CAP is NA and VA = LAB) we set CAP = 0.

```{r}
df<- df %>%
  mutate(CAP = replace(CAP,
                       VA==LAB & is.na(CAP),
                       0))
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter and store data for main industries and the aggregate economy separately

```{r}

# This list gives the codes of the main industries
main_industry_codes<- c("AtB",
                        "C",
                        "D",
                        "E",
                        "F",
                        "G",
                        "H",
                        "I",
                        "J",
                        "K",
                        "L",
                        "M",
                        "N",
                        "O",
                        "P",
                        "Q")

# filter for these codes
main_industry_accounts<- df %>%
  filter(code %in% main_industry_codes)

# Store reported totals separately

reported_totals<- df %>%
  filter(code == "TOT") %>%
  select(-code)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop country-year-industry observations that do not have all variables needed

There are some observations (n=`r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.

```{r}
kable(
  main_industry_accounts %>% filter(is.na(VA) | is.na(CAP) | is.na(LAB)) %>% count(country, code)
)

We filter these out.

```{r}

  # TODO check if COMP is needed, and add in check here and/or calculate it where missing? if so)

main_industry_accounts<- main_industry_accounts %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Add industry code descriptions

Here are the industry descriptions provided in the data.
```{r}

industry_descriptions<- df_orig %>%
  filter(code %in% main_industry_codes) %>%
  select(code, desc) %>%
  drop_na() %>%
  unique() %>%
  arrange(code)

kable(industry_descriptions)

We join these back into the clean dataframe.
```{r}

main_industry_accounts<- left_join(main_industry_accounts, industry_descriptions)

```
:::

## Sense checking

### How does the sum across all industries compare to the reported economy-wide totals?

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate aggregate across industries and merge with reported economy-wide totals

```{r}
agg_industries<- main_industry_accounts %>%
    group_by(country, year) %>%
    summarise(VA = sum(VA, na.rm = TRUE),
              CAP = sum(CAP, na.rm = TRUE),
              COMP = sum(COMP, na.rm = TRUE),
              LAB = sum(LAB, na.rm = TRUE)
              )

reported_totals <- reported_totals %>%
  rename(VA_reported = VA,
         CAP_reported = CAP,
         COMP_reported = COMP,
         LAB_reported = LAB)

check_totals<- left_join(agg_industries, reported_totals) %>%
  mutate(VA_ratio = VA/VA_reported,
         CAP_ratio = CAP/CAP_reported,
         COMP_ratio = COMP/COMP_reported,
         LAB_ratio = LAB/LAB_reported)

tolerance<- 0.001 # tenth of a percent tolerance

```
:::

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.

Note that in Canada - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, Capital income, Labour income and Compensation of employees) by around 5%, with no trend over time.

::: {.panel-tabset}
## Value added
```{r}

kable(
  check_totals %>%
  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
    select(country, year, VA_ratio, VA, VA_reported)
)
```
## Capital income
```{r}
kable(
  check_totals %>%
  filter(CAP_ratio<(1-tolerance) | CAP_ratio>(1+tolerance)) %>%
    select(country, year, CAP_ratio, CAP, CAP_reported)
)

## Labour income

```{r}
kable(
  check_totals %>%
  filter(LAB_ratio<(1-tolerance) | LAB_ratio>(1+tolerance)) %>%
    select(country, year, LAB_ratio, LAB, LAB_reported)
)

## Comp. of employees
```{r}
kable(
  check_totals %>%
  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
    select(country, year, COMP_ratio, COMP, COMP_reported)
)

:::

## Save cleaned data

::: {.callout-note collapse="true" appearance="minimal"}
### Write cleaned data to csv
```{r}
fp<- "PhD_pages/data/clean/"

write.csv(main_industry_accounts, paste0(fp, "KLEMS_gen_income_by_industry", ".csv"))

:::

### 2022-11-30 PhD_pages/data_appendices/3_B_prepare_OECD.qmd
Updating papers and data appendices

---
title: "Appendix 3.B: Preparing OECD data"
format: html
---

```{r}
library(tidyverse)
library(rsdmx)
```

## Queries of the OECD API

```{r}
start_year<- 1950
end_year<- 2020

@@

### Industry-level data on value added and its components

@@

select_table<- "SNA_TABLE6A_ARCHIVE" # Value added and its components by activity, ISIC rev4

countries<- "" # leave blank for all entities
transactions<- "" # i.e. output/GOS/CoE etc. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "C+VOB" # Prices. C = constant; VOB = indexed to OECD base year

key_query<- paste(countries,transactions,activities,measures, sep = ".")

industry_accounts <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

industry_accounts<- as.data.frame(industry_accounts)

# Save
save(industry_accounts, file="PhD_pages/data/original/OECD/industry_accounts.Rda")

@@

@@

# GDP AT PPPs ----

GDP_PPP<- readSDMX(providerId = "OECD",
                       resource = "data",
                      flowRef = "SNA_TABLE1",
                      key = ".B1_GE.VPVOB",
                       start = 1950,
                      end = 2020,
                      dsd = TRUE)

GDP_PPP<- as.data.frame(GDP_PPP)

# Save
save(GDP_PPP, file="PhD_pages/data/original/OECD/GDP_PPP_raw.Rda")

# HOUSING CONSUMPTION ------
rental_consumption <- readSDMX(providerId = "OECD",
                            resource = "data",
                            flowRef = "SNA_TABLE5",
                            key = ".P31CP041+P31CP042.C+VOB",
                            start = 1950,
                            end = 2020,
                            dsd = TRUE)

rental_consumption<- as.data.frame(rental_consumption)

# Save
save(rental_consumption, file="PhD_pages/data/original/OECD/rental_consumption.Rda")

# RENT PRICE -----
# This is the CPI component for actual rentals or else the aggrgeate
  # of actual and imputed rentals and maintenance. See paper for details.

rent_price_index <- readSDMX(providerId = "OECD",
                               resource = "data",
                               flowRef = "HOUSE_PRICES",
                               key = ".RPI",
                               start = 1950,
                               end = 2020,
                               dsd = TRUE)

rent_price_index<- as.data.frame(rent_price_index)

# Save
save(rent_price_index, file="PhD_pages/data/original/OECD/rent_price_index.Rda")

# CPI ------

CPI <- readSDMX(providerId = "OECD",
                             resource = "data",
                             flowRef = "PRICES_CPI",
                             key = ".CPALTT01.IXOB.A",
                             start = 1950,
                             end = 2020,
                             dsd = TRUE)

CPI<- as.data.frame(CPI)

  # the single subject-measure CPALTT01 - IXOB refers to all items
    # CPI index. 2015=100

# Save
save(CPI, file="PhD_pages/data/original/OECD/CPI.Rda")

# GDP deflator ------

GDP_deflator <- readSDMX(providerId = "OECD",
                resource = "data",
                flowRef = "EO",
                key = ".PGDP.A",
                start = 1950,
                end = 2020,
                dsd = TRUE)

GDP_deflator<- as.data.frame(GDP_deflator)

# the single subject-measure CPALTT01 - IXOB refers to all items
# CPI index. 2015=100

# Save
save(GDP_deflator, file="PhD_pages/data/original/OECD/GDP_deflator.Rda")

# SECTOR ACCOUNTS ------

# Grab full dataset

sector_accounts <- readSDMX(providerId = "OECD",
                        resource = "data",
                        flowRef = "SNA_TABLE14A",
                        key = "..S1+S2+S11+S12+S13+S14_S15+S14+S15.C",
                        start = 1950,
                        end = 2020,
                        dsd = TRUE)

sector_accounts<- as.data.frame(sector_accounts)

# Save
save(sector_accounts, file="PhD_pages/data/original/OECD/sector_accounts.Rda")

# INDUSTRY LABOUR INPUTS DATA ----

# set paramaters for API query

select_table<- "SNA_TABLE7A_ARCHIVE" # Labour intput by activity, ISIC rev4

start_year<- 1950
end_year<- 2020

countries<- "" # leave blank for all entities
transactions<- "" # i.e. total employment/employed/self-employed. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "HRS" # i.e. hours/jobs/fte. HRS = hours in millions

key_query<- paste(countries,transactions,activities,measures, sep = ".")

employment_by_industry <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

employment_by_industry<- as.data.frame(employment_by_industry)

# Save
save(employment_by_industry, file="PhD_pages/data/original/OECD/employment_by_industry.Rda")

@@

>>>>

---
title: "Appendix 3.B: Preparing OECD industry data"
format: html
---

## Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the OECD industry data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Load libraries and functions

```{r}
library(tidyverse)
library(rsdmx)
library(knitr)

source("PhD_pages/scripts/general_functions.R")
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load industry generation of income and employment data

```{r}
fp<- "PhD_pages/data/original/OECD/"

load(paste0(fp, "industry_accounts", ".Rda"))
load(paste0(fp, "employment_by_industry", ".Rda"))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Standardize country names

Here is the mapping of country names we use to standardize across data sources.
```{r}

OECD_country_mapping<- read.csv(paste0(fp, "OECD_country_name_mapping", ".csv"))

kable(OECD_country_mapping)

@@

Here we map the two data sets to these standardized names.
```{r}

industry_accounts<- map_names(
  df = industry_accounts, 
  df_mapping = OECD_country_mapping, 
  old_country_varname_df = "LOCATION", 
  old_country_varname_df_mapping = "OECD_country_name",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

employment_by_industry<- map_names(
  df = employment_by_industry, 
  df_mapping = OECD_country_mapping, 
  old_country_varname_df = "LOCATION", 
  old_country_varname_df_mapping = "OECD_country_name",
  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
  new_country_varname_final = "country"
)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Clean year variable in both files

```{r}
clean_year<- function(df){
  df<- df %>%
    mutate(year = as.numeric(obsTime))
}

industry_accounts<- clean_year(industry_accounts)
employment_by_industry<- clean_year(employment_by_industry)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Further tidying of industry generation of income data

@@

# Filter for only current price data and the needed transactions
industry_accounts<- industry_accounts %>%
  filter(MEASURE == "C" & 
           TRANSACT %in% c("B1GA", "D1A","B2G_B3GA", "B2N_B3NA", "D29_D39A"))

# Recode transactions from national accounts codes to something more legible
recoding_transactions <- c(
  B1GA = "VA", # (Gross) Value added
  D1A = "COMP", # Compensation of employees
  B2G_B3GA = "GOS", # Gross operating surplus
  B2N_B3NA = "NOS", # Net operating surplus (net of depreciation)
  D29_D39A = "TOP") # Taxes on production

industry_accounts<- industry_accounts %>%
  mutate(TRANSACT = recode(TRANSACT, !!!recoding_transactions))

# flip to wide format
industry_accounts<- industry_accounts %>%
  pivot_wider(names_from = TRANSACT, values_from = obsValue)

# Select and rename needed vars
industry_accounts<- industry_accounts %>%
  rename(ind_code = ACTIVITY) %>%
  select(country, year, ind_code, VA, COMP, GOS, NOS, TOP)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter and store generation of income data for main industries and the aggregate economy separately

# Main aggregate industries

main_industry_codes<- c("VA0", "VB", "VC", "VD", "VE", "VF", "VG", "VH", "VI",
                        "VJ", "VK", "VL", "VM", "VN", "VO", "VP", "VQ", "VR",
                        "VS", "VT", "VU")

# Filter industry data for main industry codes only
main_industry_accounts<- industry_accounts %>%
  filter(ind_code %in% main_industry_codes)

# Store economy-wide aggregates separately
reported_totals <- industry_accounts %>%
  filter(ind_code == 'VTOT')

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Further tidying of industry employment data.

```{r}
# Select self-employed and employed hours and flip to wide format
employment_by_industry<- employment_by_industry %>%
  filter(TRANSACT %in% c("ESEA", "EEMA")) %>%  # codes for self-employed and employed
  pivot_wider(id_cols = c(country, year, ACTIVITY), names_from = TRANSACT, values_from = obsValue)

# Select and rename needed vars
employment_by_industry<- employment_by_industry %>%
  rename(self_employed_hrs = ESEA, 
         employed_hrs = EEMA,
         ind_code = ACTIVITY)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Join generation of income and employment data

```{r}
main_industry_accounts<- left_join(main_industry_accounts, employment_by_industry)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter out observations where the implied wage is infinite

In the joined data, there are a small number of rows where employed hours = 0 but compensation of employees \> 0 --- implying an infinite wage.

Here we show these illogical rows, grouped by country and industry.

```{r}
kable(
  main_industry_accounts %>%
  filter((COMP>0 & employed_hrs == 0))
)
```

We drop these.

```{r}
main_industry_accounts<- main_industry_accounts %>%
  filter(!(COMP>0 & employed_hrs == 0))

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Filter out country-year-industry observations without all the variables we need

There are some rows where we do not have all the variables we will need for the calculations: VA, COMP, GOS, and employed and self-employed hours.

```{r}
kable(
  main_industry_accounts %>%
    filter(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)) %>%
    count(country, ind_code)
)
```

We drop these.

```{r}
main_industry_accounts<-  main_industry_accounts %>%
    filter(!(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)))
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
#### Filter out country-year observations with partial industry coverage

The objective is to have complete coverage of the economy. There are some country years where coverage across industries looks decidedly incomplete.

For instance, out of `r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:

```{r}
kable( 
 to_drop<- main_industry_accounts %>%
   count(country, year) %>%
   filter(n<18)
)

@@

We drop these country-years with apparently incomplete data.

```{r}
main_industry_accounts<- left_join(main_industry_accounts, to_drop) %>%
  filter(is.na(n)) %>%
  select(-n)

But because of differences in national accounting methods, missing data for a given industry does not imply the accounts are incomplete.

In particular, the dataset contains only `r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to `r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.

```{r}
summary(main_industry_accounts %>% filter(ind_code == "VU"))
```

In the table below we show all remaining country years in which at least one industry --- other than `VU` --- is missing.

We assume that these apparent gaps do not indicate incompleteness of the data, rather just differences in national accounting practices.

@@

kable( main_industry_accounts %>%
  select(year, country, ind_code, VA) %>%
  pivot_wider(names_from = ind_code, values_from = VA) %>%
  filter(if_any(VA0:VT, ~ is.na(.)))
)

```
:::

## Sense checking

### How does the sum across all industries compare to the reported economy-wide totals?

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate aggregate across industries and merge with reported economy-wide totals

```{r}
agg_industries<- main_industry_accounts %>%
    group_by(country, year) %>%
    summarise(VA = sum(VA, na.rm = TRUE),
              COMP = sum(COMP, na.rm = TRUE),
              GOS = sum(GOS, na.rm = TRUE),
              NOS = sum(NOS, na.rm = TRUE)
              )

reported_totals <- reported_totals %>%
  rename(VA_reported = VA,
         COMP_reported = COMP,
         GOS_reported = GOS,
         NOS_reported = NOS) %>%
  select(-ind_code)

check_totals<- left_join(agg_industries, reported_totals) %>%
  mutate(VA_ratio = VA/VA_reported,
         COMP_ratio = COMP/COMP_reported,
         GOS_ratio = GOS/GOS_reported,
         NOS_ratio = NOS/NOS_reported)

tolerance<- 0.005 # half a percent tolerance

```
:::

In general the sum across industries is very close to to the reported economy-wide aggregate. In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.

Note that in Finland - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, compensation of employees and gross operating surplus) by around 5%, with no trend over time.

::: {.panel-tabset}
## Value added
```{r}

kable(
  check_totals %>%
  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
    select(country, year, VA_ratio, VA, VA_reported)
)
```
## Comp. of employees
```{r}
kable(
  check_totals %>%
  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
    select(country, year, COMP_ratio, COMP, COMP_reported)
)

```
## Gross operating surplus
```{r}
kable(
  check_totals %>%
  filter(GOS_ratio<(1-tolerance) | GOS_ratio>(1+tolerance)) %>%
    select(country, year, GOS_ratio, GOS, GOS_reported)
)

@@

:::

## Save cleaned data

::: {.callout-note collapse="true" appearance="minimal"}
### Write cleaned data to csv
```{r}
fp<- "PhD_pages/data/clean/"

write.csv(main_industry_accounts, paste0(fp, "OECD_gen_income_and_employment_by_industry", ".csv"))

```
:::

### 2022-11-30 PhD_pages/paper_pages/paper3.qmd
Updating papers and data appendices

@@

What share of aggregate income is made up by rents on housing assets and how has this evolved in recent decades? The national accounts offers three avenues for answering this question: operating surplus within the household sector ('Household sector' approach); operating surplus within the real estate industry ('Real estate industry' approach); and household final consumption of housing services ('Final consumption' approach).

Figure 1 shows the housing capital share of factor cost value added calculated according to three approaches.

>>>>

## Introduction

A substantial literature has documented a long-run shift in the distribution of national income between wages and returns to capital.

Much of the interest in this empirical trend owes to its lying at the junction of two prominent topics within the social sciences: 1) the decline of labour bargaining power, and 2) the rise in income inequality.

(Explain (1) and connection to (2)).

The hypothesis underpins much of the motivation for the Top incomes approach.

## Data

###  Background on the structure of national accounts

### Approaches to calculating factor shares

### Available sources

@@

The unusual nature of housing ... in the national accounts.

As a *stock*, housing is a non-financial asset included in the balance sheet. Housing also features as a *flow*, included in each of the three approaches to calculating GDP: output, income and expenditure.

As output, housing services are included in Real Estate activity (SIC....). Sub Sector.

This flow of housing services generates a stream of income, which, in addition to small amounts of labour income (property management, agents' fees), is the yield received by owners of housing assets.[^1] This private or public rented sector , and owner occupied, to which an equivalent rent is imputed. includes rent imputed to owner occupiers, as well as actual rents received by private landlords, whether individuals or corporations (both privately and publicly owned).

[^1]: ADD NOTE ABOUT LABOUR INCOME IN ACTUAL RENTS

This income as an expenditure by households, again both actual and imputed rents 'paid'.

In line with national accounting principles, the value of the flow of housing is equal, whether conceived of as a stream of services, income, or expenditure. As such, these offer different entry points from which aggregate housing income can be calculated and isolated from non-housing capital income.

Figure 1 shows these housing flows as a share of total income.

In each case the denoninator is gross domestic value added, valued at factor cost. This aggregate is similar to GDP, with the exception that taxes and subsidies on production are excluded. This results in an aggregate that capital and labour components: *gross operating surplus* and *compensation of employees.*
